---
title: "Assignment B-1: Making a function"
author: "Irvin Ng"
date: "2023-11-03"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(testthat))
```


# Function Definition
I wanted to make a function which gives me relevant summary data. In particular, I wanted the function to give me the counts, mean, median, and SD (where applicable). Even if the data was not numeric, it would at least give me a count of all the non-numeric data (eg. the amount of countries in a df).

```{r}
#' @Title Generate summary statistics
#'
#' This function calculates the different summary statistics (eg.count, mean, median, and standard deviation) of numeric values
#' in the specified column of a data frame. In the case of non-numeric values, it would at least get counts.
#'
#' @param data A data frame containing the column to be summarized (e.g., gapminder dataset). I named this data because it seems to be a 
#' very common name for the object even for other functions in other packages.
#' @param column_name The name of the column in the data frame to be summarized.
#' @return The count, mean, median, and standard deviation of numeric values.

numeric_summary <- function(data, column_name) {
  # Check if the input data is a data frame
  if (!is.data.frame(data)) {
    stop("Input must be a data frame.")
  }
  
  # Check if the input column exists in the data frame
  if (!column_name %in% names(data)) {
    stop("Specified column does not exist in the data frame.")
  }
  
  # Extract numeric values from the specified column, excluding NA values
  numeric_values <- na.omit(data[[column_name]])
  
  # Check if the column contains numeric data
  if (length(numeric_values) == 0) {
    stop("Specified column does not contain numeric data.")
  }
  
  # Calculate summary statistics with NA values removed
  summary_stats <- list(
    Count = length(numeric_values),
    Mean = mean(numeric_values, na.rm = TRUE),
    Median = median(numeric_values, na.rm = TRUE),
    Standard_Deviation = sd(numeric_values, na.rm = TRUE)
  )
  
  # Return the summary statistics
  return(summary_stats)
}

gapminder

```


# Examples
I will be using the gapminder dataset to demonstrate my function. Here, we will look at the population, GDP per capita, and the Country (non-numeric)

## Example 1: Generate summary for the 'pop' column in the gapminder dataset
```{r}
numeric_summary(gapminder, "pop")
```

## Example 2: Generate summary for the 'gdpPercap' column in the gapminder dataset
```{r}
numeric_summary(gapminder, "gdpPercap")
```

## Example 3: Generate a summary for a non-numeric column (we'll see errors)
```{r error=TRUE}
numeric_summary(gapminder, "country")
```

# Error Testing

```{r}
# Make a custom df with NA values to do the second test:
NA_df <- gapminder
set.seed(1) # Set seed for reproducibility
NA_df$gdpPercap[sample(1:nrow(NA_df), 10)] <- NA # Add NA values to 10 random rows


test_that("numeric_summary function works as expected", {
  # Test 1: Check if the function works with a numeric column without NA values
  expect_equal(numeric_summary(gapminder, "pop")$Count, 1704)
  
  # Test 2: Check if the function works with a numeric column with NA values
  expect_true(!is.na(numeric_summary(NA_df, "gdpPercap")$Mean))
  
  # Test 3: Check if the function handles non-numeric columns and throws an error
  expect_error(numeric_summary(gapminder, "country"))
})


```

